---
title: "Finding Question Notes"
output:
  html_document: default
vignette: >
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteIndexEntry{Finding Question Information}
  %\usepackage[utf8]{inputenc}
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This document answers the following questions through the example of question notes:
"How do I find specific information or properties about questions from a QSF file?"
"How do I find which questions have specific information or properties in their structure?

This document demonstrates the use of for loops, the *apply family of functions, 
and some data type conversion and constructions with lists, vectors, and dataframes.

Finally, this document answers one more question: How do I view dataframes with large amounts
of text easily in R?

```{r}
# Load the local version of the package
devtools::load_all("~/Documents/DataAnalysis/QualtricsTools_fork")
# Load an example survey with notes.
get_setup(headerrows=3, 
          qsf_path = file.path(
            path.package("QualtricsTools"),
            "/data/Sample Surveys/User Notes Survey/Notes_Survey.qsf"
            ),
          csv_path = file.path(
            path.package("QualtricsTools"),
            "/data/Sample Surveys/User Notes Survey/Notes_Survey.csv"
            )
)
```

First, how do we find questions which have user notes?

When we use get_setup, this calls get_coded_questions_and_blocks, which
inserts notes into questions as list elements of question[['qtNotes']] with
"User Note: " prepended.

This just prints the questions' data export tags to the screen for
each question that has user notes added.

```{r}
questions_with_notes <- list()  # Make an empty list for storing DataExportTags
for (question in questions) {
  if ('qtNotes' %in% names(question)) { # Check that question[['qtNotes']] exists
    # user_notes_tf is a list of TRUE/FALSE values for each qtNote that indicates
    # whether or not "User Note: " is a substring of the qtNote.
    user_notes_tf <- grepl("User Note: ", question[['qtNotes']])
    # If there are any qtNotes which are user notes as indicated by user_notes_tf
    # for the current question, append the current question's dataexporttag to the
    # list of questions_with_notes.
    if (any(user_notes_tf)) {
      questions_with_notes <- c(questions_with_notes,
                                question[['Payload']][['DataExportTag']])
    }
  }
}
print(questions_with_notes)
```


There are two ways that one could make the above code into a function:
By simply wrapping this with something like
```{r}
find_user_notes_questions <- function(questions) {
...
}
```
Or, we could turn the inside logic which determines whether or not
a question has user notes into a function that takes an argument
of a question which we can apply to the entire list of questions.
This is a more modular approach, and is preferable because then we
can use the "has_user_notes(question)" function in more ways.
Here I will demonstrate the second of these ways.


```{r}
has_user_notes <- function(question) {
  if ('qtNotes' %in% names(question)) { # Check that question[['qtNotes']] exists
    # user_notes_tf is a list of TRUE/FALSE values for each qtNote that indicates
    # whether or not "User Note: " is a substring of the qtNote.
    user_notes_tf <- grepl("User Note: ", question[['qtNotes']])
    # If there are any qtNotes which are user notes as indicated by user_notes_tf
    # for the current question, append the current question's dataexporttag to the
    # list of questions_with_notes.
    if (any(user_notes_tf)) {
      return(TRUE)
    }
  }
  # If the previous return wasn't executed, then the question either has no qtNotes
  # or no notes with "User Notes: " as a substring.
  return(FALSE)
}
```

Now we use sapply to run the has_user_notes function on every
question and return a logical vector with TRUE/FALSE values for
each question. We're going to use sapply as opposed to lapply
because lapply always returns a list structure (which is
not necessarily of homogenous datatype, i.e. list(TRUE, "a") is a
valid list in R) whereas sapply
simplifies the output to a homogenous datatype such as vector or
matrix where possible. (vectors and matrices are "atomic" data structures in
R, meaning that they are homogenous in the sense that
all their entries must be of the same datatype so that the vector is
strictly one of "logical", "integer", "numeric", "double", "complex",
"character" or "raw".
```{R}
sapply(questions, has_user_notes)
```

Now that we have a logical vector, we can find out which indices of questions
have user notes.
```{r}
questions_with_notes_indices <- which(sapply(questions, has_user_notes))
```

To just get the dataexporttags of these questions, something like this
is effective:
```{r}
dataexporttags_with_notes <- sapply(questions_with_notes_indices, function(i) {
  questions[[i]][['Payload']][['DataExportTag']]
  })
```

If we want a list of these questions, we can use list subsetting:
```{r}
questions_with_notes <- questions[questions_with_notes_indices]
```

Using str to see the structure of questions_with_notes, we can
see that it is in fact a list of three questions that can be
verified to have data export tags.
```{r}
str(questions_with_notes, max.level = 1)
```


This function allows us to input a specific question and get back a list of the user notes.
This uses gsub to remove the prepended "User Note: " portion of the user notes.
```{r}
list_user_notes <- function(question) {
  if ('qtNotes' %in% names(question)) { # Check that question[['qtNotes']] exists
    # user_notes_tf is a list of TRUE/FALSE values for each qtNote that indicates
    # whether or not "User Note: " is a substring of the qtNote.
    user_notes_tf <- grepl("User Note: ", question[['qtNotes']])
    # If there are any qtNotes which are user notes as indicated by user_notes_tf
    # for the current question, append the current question's dataexporttag to the
    # list of questions_with_notes.
    if (any(user_notes_tf)) {
      user_notes <- question[['qtNotes']][which(user_notes_tf)]
      cleaned_user_notes <- sapply(user_notes, function(x) gsub("User Note: ", "", x))
      return(cleaned_user_notes)
    }
  }
}
list_user_notes(questions[[1]])
```

We can pair the output information from the list_user_notes with the data export tag in
order to be able to create a character matrix which shows for each data export tags what
user notes are associated to it. The data export tags will appear in the first row,
and the notes will appear in the second.
```{r}
table_question_notes <- function(question) {
  sapply(list_user_notes(question), function(x) {
    c(question[['Payload']][['DataExportTag']], x)
    })
}
table_question_notes(questions[[1]])

table_all_notes <- function(questions) sapply(questions, table_question_notes)
all_notes <- table_all_notes(questions)
all_notes
```

Notice a few things about table_all_notes:
It outputs a list, which is not yet a dataframe as we would like the end result to be.
If a question does not have any notes, then for the ith question it will include an empty 
`list()` in its output (where i represents an actual integer).
We can replicate this scenario by just removing the qtNotes from questions[[2]].
```{r}
questions[[2]][['qtNotes']] <- NULL
all_notes <- table_all_notes(questions)
all_notes
```

To remove the empty list() elements in , we need a delete.NULLs function which removes any empty sublists.
The unlist function returns a flat atomic vector (coercing entries into the same datatype if need-be) which
we can use to find list elements which have no contents, i.e. length 0. We then subset including only elements
which when unlisted have length not 0 and return the sublist.
```{r}
delete.NULLs  <-  function(x){
  x[unlist(lapply(x, length) != 0)]
}
```

We can use all_notes to delete the empty list in [[2]].
```{r}
all_notes <- delete.NULLs(all_notes)
all_notes
```

Now we can actually turn this list into a useable dataframe by simply using as.data.frame:
```{r}
notes_df <- as.data.frame(all_notes)
notes_df
```

Let's get rid of those unnecessary and unhelpful column names.
```{r}
colnames(notes_df) <- NULL
notes_df
```



### Conclusion

I hope this document was helpful in explaining how to find and select 
questions based on particular criterion, how to find specific information 
inside questions, and how to view the data you find more easily.
Further directions I would suggest include learning about the 
[apply family](https://www.datacamp.com/community/tutorials/r-tutorial-apply-family)
in the base-R package and about
[iteration in Hadley Wickham's R for Data Science book](http://r4ds.had.co.nz/iteration.html).
  
